---
title: "Untitled"
author: "Chloe Child"
date: "2025-07-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(forcats)
library(gamlss)
library(ggpubr)
```


```{r}
getwd()
bayley_composite <- readRDS("Bayley_Composite_scores.rds")

bayley_gsv <- readRDS("Bayley_GSV_scores.rds")

bayley_scaled <- readRDS("Bayley_Scaled_scores.rds")

head(bayley_composite)
```
## Convert bayley_composite to long form

```{r}
composite_rename <- bayley_composite %>% rename(
  Cognitive = 'bsid_cog_composite',
  Language = 'bsid_lang_composite',
  Motor = 'bsid_mot_composite',
)

head(composite_rename)
composite_long <- composite_rename %>% pivot_longer(cols=c('Cognitive', 'Language', 'Motor'), names_to = 'domain', values_to = 'score')

head(composite_long)
```

## Convert bayley_gsv to long form

```{r}
head(bayley_gsv)
gsv_rename <- bayley_gsv %>% rename(
  Cognitive = 'bsid_gsv_cog',
  Receptive = 'bsid_gsv_rc',
  Expressive = 'bsid_gsv_ec',
  `Fine Motor` = 'bsid_gsv_fm',
  `Gross Motor` = 'bsid_gsv_gm'
)



gsv_long <- gsv_rename %>% pivot_longer(cols = c('Cognitive', 'Receptive', 'Expressive', 'Fine Motor', 'Gross Motor'), names_to = 'domain', values_to = 'score')

head(gsv_long)
```

# Convert bayley_scaled to long form
```{r}
head(bayley_scaled)

scaled_rename <- bayley_scaled %>% rename(
  Cognitive = "bsid_cog_scaled",
  Receptive = "bsid_rc_scaled",
  Expressive = "bsid_ec_scaled",
  Fine_Motor = "bsid_fm_scaled",
  Gross_Motor = "bsid_gm_scaled"
)

scaled_long <- scaled_rename %>% pivot_longer(cols = c('Cognitive', 'Receptive', 'Expressive', 'Fine_Motor', 'Gross_Motor'), names_to = 'domain', values_to = 'score')

head(scaled_long)
```

# Convert Bayley GSV Scores
```{r}
## Create a function to transform scores

transform_score <- function(scores) {
  z_scores <- (scores - 500)/100
  
  transformed <- z_scores * 25 + 500
  
  return(transformed)
}
```

```{r}
new_gsv_long <- gsv_long %>% mutate(transformed_score = round(transform_score(score)),0) # Add transformed scores as a new column

head(gsv_long)
head(new_gsv_long)
```

# Remove 2 month visits from the GSV scores
```{r}
new_gsv_long_rem <- new_gsv_long %>% filter(redcap_event_name != "2_month_visit_arm_1")

table(new_gsv_long_rem$redcap_event_name)
```


```{r}
summary(new_gsv_long_rem)
```

```{r}
ggplot(new_gsv_long_rem, aes(x = bsid_age_calc, y = transformed_score)) + geom_smooth(aes(color = as.factor(sca_condition))) + geom_point(x = new_gsv_long_rem$bsid_age_calc, y = new_gsv_long_rem$transformed_score, alpha = 0.5, color = "#4371c1") + facet_wrap(~domain)
```

```{r}
ggplot(new_gsv_long_rem, aes(x = bsid_age_calc, 
                             y = fct_reorder(as.factor(domain), bsid_age_calc), 
                             )) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +  # Hide default outliers
  geom_jitter(aes(color = transformed_score), 
              width = 0.2, height = 0.1, size = 2, alpha = 0.9) +  # Show transformed_score

  scale_color_gradient(low = "skyblue", high = "darkblue") +  # Color scale for transformed_score
  labs(
    x = "Age Score was Achieved (Months)",
    y = NULL,
    fill = "SCA Condition",
    color = "Transformed Score"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 14),
    legend.title = element_text(size = 14)
  ) +facet_wrap(~as.factor(sca_condition))


```

```{r}
ggplot(new_gsv_long_rem, aes(x = transformed_score, 
                             y = fct_reorder(as.factor(domain), transformed_score), 
                             fill = sca_condition)) +
  geom_violin(trim = FALSE, alpha = 0.5) +
  geom_jitter(aes(color = transformed_score), width = 0.1, height = 0.2, size = 1.5, alpha = 0.7) +
  scale_color_gradient(low = "skyblue", high = "darkblue") +
  labs(
    x = "Transformed Score",
    y = NULL,
    fill = "SCA Condition",
    color = "Transformed Score"
  ) +
  theme_minimal(base_size = 16)

```


```{r}
library(ggridges)

ggplot(new_gsv_long_rem, aes(x = transformed_score, 
                             y = fct_reorder(domain, transformed_score), 
                             fill = sca_condition)) +
  geom_density_ridges(alpha = 0.6, scale = 1.2, rel_min_height = 0.01) + facet_wrap(~as.factor(redcap_event_name)) + 
  labs(
    x = "Transformed Score",
    y = NULL,
    fill = "SCA Condition"
  ) +
  theme_minimal(base_size = 16)
```

```{r}
ggplot(new_gsv_long_rem, aes(x = transformed_score, 
                             y = fct_reorder(domain, transformed_score))) +
  geom_density_ridges(aes(fill = sca_condition), alpha = 0.4, scale = 1.1) +
  geom_jitter(aes(color = bsid_age_calc), width = 0.1, height = 0.1, alpha = 0.6, size = 1.5) +

  scale_color_gradient(low = "orange", high = "red") +
  labs(
    x = "Transformed Score",
    y = NULL,
    fill = "SCA Condition",
    color = "Age (Months)"
  ) +
  theme_minimal(base_size = 16)

```
# Composite Violin Updated
```{r}
ref_score_min <- 100 - 15  # 85
ref_score_max <- 100 + 15  # 115

ymin_val <- min(composite_long$bsid_age_calc, na.rm = TRUE)
ymax_val <- max(composite_long$bsid_age_calc, na.rm = TRUE)



# Plot
violin_plot <- ggplot(composite_long, aes(
  x = fct_reorder(sca_condition, bsid_age_calc),
  y = bsid_age_calc,
  fill = sca_condition
)) +
  geom_violin(trim = FALSE, alpha = 0.5, color = NA) +
  geom_jitter(aes(color = score),
              width = 0.2, height = 0.1, size = 2, alpha = 0.5) +
  scale_color_gradient(low = "lightblue", high = "darkorange") +
  facet_wrap(~domain, ncol = 5) +
  coord_flip() +
  labs(
    x = NULL,
    y = "Age at Assessment (Months)",
    fill = "SCA Condition",
    color = "Score"
  ) +
  theme_bw(base_size = 16) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )


violin_plot +
  geom_rect(
    inherit.aes = FALSE,
    aes(xmin = ref_score_min, xmax = ref_score_max,
        ymin = ymin_val, ymax = ymax_val),
    fill = "lightblue", alpha = 0.2
  )

```


## Violin Graphs with BoxPlot

```{r}

# Normative (reference) box values for score
comp_ref_mean <- 100
comp_ref_sd <- 15

# Reference boxplot data (use same structure as violin groups)
ref_box_data_composite <- data.frame(
  group = unique(composite_long$sca_condition),
  ymin = comp_ref_mean - comp_ref_sd,
  lower = comp_ref_mean - comp_ref_sd / 2,
  middle = comp_ref_mean,
  upper = comp_ref_mean + comp_ref_sd / 2,
  ymax = comp_ref_mean + comp_ref_sd
)

# Corrected plot
violin_plot_comp <- ggplot(composite_long_filtered, aes(
  x = fct_reorder(sca_condition, score),
  y = score,
  fill = sca_condition
)) +  geom_violin(trim = FALSE, alpha = 0.5, color = NA) +
  geom_jitter(aes(color = bsid_age_calc), 
              width = 0.2, height = 0.1, size = 2, alpha = 0.7) +
  geom_boxplot(
    data = ref_box_data_composite,
    aes(
      x = group,
      ymin = ymin,
      lower = lower,
      middle = middle,
      upper = upper,
      ymax = ymax
    ),
    stat = "identity",
    inherit.aes = FALSE,
    width = 0.2,
    fill = "orange",
    color = "black",
    alpha = 0.5) +
 scale_color_gradient(low = "skyblue", high = "navy") +
  facet_wrap(~domain, ncol = 5) +
  labs(
    x = NULL,
    y = "Score",
    fill = "SCA Condition",
    color = "Age at Assessment"
  ) +
  theme_bw(base_size = 16) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
      legend.text = element_text(size = 8),
  legend.title = element_text(size = 10)
    
  )

violin_plot_comp

ggexport(violin_plot_comp, filename = "C:/Users/craft/OneDrive/Documents/GitHub/eXtraordinarYBabiesStudy_ShinyApps/Chloe/Bayley Progression Code/Bayley_take1/bayleytab1/comp_plot_violin.png", width = 6000, height = 3000, res = 250)

```

## Violin Plot with Boxplot for Scaled
```{r}

scaled_long_filtered <- scaled_long %>% filter(score < 777)

# Normative (reference) box values for score
scale_ref_mean <- 10
scale_ref_sd <- 3

# Reference boxplot data (use same structure as violin groups)
ref_box_data_scaled <- data.frame(
  group = unique(scaled_long$sca_condition),
  ymin = scale_ref_mean - scale_ref_sd,
  lower = scale_ref_mean - scale_ref_sd / 2,
  middle = scale_ref_mean,
  upper = scale_ref_mean + scale_ref_sd / 2,
  ymax = scale_ref_mean + scale_ref_sd
)

# Corrected plot
violin_plot_scaled <- ggplot(scaled_long_filtered, aes(
  x = fct_reorder(sca_condition, score),
  y = score,
  fill = sca_condition
)) +
  geom_violin(trim = FALSE, alpha = 0.5, color = NA) +
  geom_jitter(aes(color = bsid_age_calc),  # optional: color by age at assessment
              width = 0.2, height = 0.1, size = 2, alpha = 0.7) +
  geom_boxplot(
    data = ref_box_data_scaled,
    aes(
      x = group,
      ymin = ymin,
      lower = lower,
      middle = middle,
      upper = upper,
      ymax = ymax
    ),
    stat = "identity",
    inherit.aes = FALSE,
    width = 0.2,
    fill = "orange",
    color = "black",
    alpha = 0.5) +
  # scale_color_gradient(low = "skyblue", high = "navy") +
  facet_wrap(~domain, ncol = 5) +
  labs(
    x = NULL,
    y = "Score",
    fill = "SCA Condition",
    color = "Age at Assessment"
  ) +
  theme_bw(base_size = 16) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
          legend.text = element_text(size = 8),
  legend.title = element_text(size = 10)
    
  )

violin_plot_scaled

ggexport(violin_plot_scaled, filename = "C:/Users/craft/OneDrive/Documents/GitHub/eXtraordinarYBabiesStudy_ShinyApps/Chloe/Bayley Progression Code/Bayley_take1/bayleytab1/violin_plot_scaled.png", width = 2100, height = 1800, res = 250)
```

## Violin Plot with Boxplot for GSV
```{r}


# Normative (reference) box values for score
gsv_ref_mean <- 500
gsv_ref_sd <- 25

# Reference boxplot data (use same structure as violin groups)
ref_box_data_gsv <- data.frame(
  group = unique(scaled_long$sca_condition),
  ymin = gsv_ref_mean - gsv_ref_sd,
  lower = gsv_ref_mean - gsv_ref_sd / 2,
  middle = gsv_ref_mean,
  upper = gsv_ref_mean + gsv_ref_sd / 2,
  ymax = gsv_ref_mean + gsv_ref_sd
)

# Corrected plot
violin_plot_gsv <- ggplot(new_gsv_long_rem, aes(
  x = fct_reorder(sca_condition, score),
  y = score,
  fill = sca_condition
)) +
  geom_violin(trim = FALSE, alpha = 0.5, color = NA) +
  geom_jitter(aes(color = bsid_age_calc),  # optional: color by age at assessment
              width = 0.2, height = 0.1, size = 2, alpha = 0.7) +
  geom_boxplot(
    data = ref_box_data_gsv,
    aes(
      x = group,
      ymin = ymin,
      lower = lower,
      middle = middle,
      upper = upper,
      ymax = ymax
    ),
    stat = "identity",
    inherit.aes = FALSE,
    width = 0.2,
    fill = "orange",
    color = "black",
    alpha = 0.5) +
  scale_color_gradient(low = "skyblue", high = "navy") +
  facet_wrap(~domain, ncol = 5) +
  labs(
    x = NULL,
    y = "Score",
    fill = "SCA Condition",
    color = "Age at Assessment"
  ) +
  theme_bw(base_size = 16) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
  legend.text = element_text(size = 8),
  legend.title = element_text(size = 10)
  )

violin_plot_gsv
ggexport(violin_plot_gsv, filename = "C:/Users/craft/OneDrive/Documents/GitHub/eXtraordinarYBabiesStudy_ShinyApps/Chloe/Bayley Progression Code/Bayley_take1/bayleytab1/violin_plot_gsv.png", width = 2100, height = 1800, res = 250)
```


```{r}
mform_dt_scre <- gamlss(formula = transformed_score ~ pb(bsid_calc_age, lambda = 5),
                       sigma.formula = ~pb(bsid_calc_age),
                       nu.formula = ~1,
                       tau.formula = ~1,
                       data = new_gsv_long_rem,
                       family = BCCG())
```

## Growth Plots for Cognitive Domain
```{r}


gsv <- new_gsv_long_rem



# Filter to Cognitive domain
gsv_cog <- gsv %>%
  filter(domain == "Cognitive")

# Plot Cognitive scores by age
ggplot(gsv_cog, aes(x = bsid_age_calc, y = transformed_score)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  labs(
    title = "Cognitive Domain Scores vs. Age",
    x = "Age (months)",
    y = "Cognitive Score"
  )

# Remove NAs and preprocess for modeling
gsv_nona_cog <- gsv_cog %>%
  filter(redcap_event_name != "2_month_visit_arm_1") %>%
  filter(!is.na(bsid_age_calc), !is.na(score))

# Fit GAMLSS model
gamlss_cog <- gamlss(
  formula = transformed_score ~ pb(bsid_age_calc, lambda = 5),
  sigma.formula = ~ pb(bsid_age_calc),
  nu.formula = ~1,
  tau.formula = ~1,
  data = gsv_nona_cog,
  family = BCCG()
)



# 1. Prediction ages
ages <- seq(2, 41, by = 0.5)

# 2. Build newdata with all required covariates used in the model
newdata <- data.frame(bsid_age_calc = ages)

# 3. Get predicted distribution parameters from the model
params <- predictAll(gamlss_cog, newdata = newdata)

# 4. Calculate centiles manually using the BCCG distribution quantile function
centiles <- c(5, 10, 25, 50, 75, 90, 95)
q_vals <- sapply(centiles / 100, function(p) {
  qBCCG(p, mu = params$mu, sigma = params$sigma, nu = params$nu)
})

# 5. Create a dataframe of predicted centiles
lms_cog <- data.frame(age = ages, q_vals)
colnames(lms_cog)[-1] <- paste0("P", centiles)


## Plot it
ggplot(gsv_nona_cog, aes(x = bsid_age_calc, y = transformed_score)) + 
  geom_point()  + # Plot Cognitive scores by age
  geom_ribbon(data = lms_cog, aes(x = age, ymin = P5, ymax = P95), inherit.aes = F, alpha = 0.5)
```

## Growth Plots for Receptive Domain
```{r}

# Filter to Cognitive domain
gsv_rec <- gsv %>%
  filter(domain == "Receptive")

# Plot Cognitive scores by age
ggplot(gsv_rec, aes(x = bsid_age_calc, y = transformed_score)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  labs(
    title = "Receptive Domain Scores vs. Age",
    x = "Age (months)",
    y = "Receptive Score"
  )

# Remove NAs and preprocess for modeling
gsv_nona_rec <- gsv_rec %>%
  filter(redcap_event_name != "2_month_visit_arm_1") %>%
  filter(!is.na(bsid_age_calc), !is.na(score))

# Fit GAMLSS model
gamlss_rec <- gamlss(
  formula = transformed_score ~ pb(bsid_age_calc, lambda = 5),
  sigma.formula = ~ pb(bsid_age_calc),
  nu.formula = ~1,
  tau.formula = ~1,
  data = gsv_nona_rec,
  family = BCCG()
)



# 1. Prediction ages
ages <- seq(2, 41, by = 0.5)

# 2. Build newdata with all required covariates used in the model
newdata <- data.frame(bsid_age_calc = ages)

# 3. Get predicted distribution parameters from the model
params <- predictAll(gamlss_rec, newdata = newdata)

# 4. Calculate centiles manually using the BCCG distribution quantile function
centiles <- c(5, 10, 25, 50, 75, 90, 95)
q_vals <- sapply(centiles / 100, function(p) {
  qBCCG(p, mu = params$mu, sigma = params$sigma, nu = params$nu)
})

# 5. Create a dataframe of predicted centiles
lms_rec <- data.frame(age = ages, q_vals)
colnames(lms_rec)[-1] <- paste0("P", centiles)


## Plot it
ggplot(gsv_nona_rec, aes(x = bsid_age_calc, y = transformed_score)) + 
  geom_point() + 
  geom_smooth(data = lms_rec, aes(x = age, y = P5)) + 
  geom_smooth(data = lms_rec, aes(x = age, y = P10)) + 
  geom_smooth(data = lms_rec, aes(x = age, y = P25)) + 
  geom_smooth(data = lms_rec, aes(x = age, y = P50)) + 
  geom_smooth(data = lms_rec, aes(x = age, y = P75)) + 
  geom_smooth(data = lms_rec, aes(x = age, y = P90)) + 
  geom_smooth(data = lms_rec, aes(x = age, y = P95)) 
```

## Growth Plots for Expressive Domain
```{r}
# Filter to Cognitive domain
gsv_exp <- gsv %>%
  filter(domain == "Expressive")

# Plot Cognitive scores by age
ggplot(gsv_exp, aes(x = bsid_age_calc, y = transformed_score)) + 
  geom_point() + 
  geom_smooth(se = TRUE) +
  labs(
    title = "Expressive Domain Scores vs. Age",
    x = "Age (months)",
    y = "Expressive Score"
  )

# Remove NAs and preprocess for modeling
gsv_nona_exp <- gsv_exp %>%
  filter(redcap_event_name != "2_month_visit_arm_1") %>%
  filter(!is.na(bsid_age_calc), !is.na(score))

# Fit GAMLSS model
gamlss_exp <- gamlss(
  formula = transformed_score ~ pb(bsid_age_calc, lambda = 5),
  sigma.formula = ~ pb(bsid_age_calc),
  nu.formula = ~1,
  tau.formula = ~1,
  data = gsv_nona_exp,
  family = BCCG()
)



# 1. Prediction ages
ages <- seq(2, 41, by = 0.5)

# 2. Build newdata with all required covariates used in the model
newdata <- data.frame(bsid_age_calc = ages)

# 3. Get predicted distribution parameters from the model
params <- predictAll(gamlss_cog, newdata = newdata)

# 4. Calculate centiles manually using the BCCG distribution quantile function
centiles <- c(5, 10, 25, 50, 75, 90, 95)
q_vals <- sapply(centiles / 100, function(p) {
  qBCCG(p, mu = params$mu, sigma = params$sigma, nu = params$nu)
})

# 5. Create a dataframe of predicted centiles
lms_exp <- data.frame(age = ages, q_vals)
colnames(lms_exp)[-1] <- paste0("P", centiles)


## Plot it
ggplot(gsv_nona_exp, aes(x = bsid_age_calc, y = transformed_score)) + 
  geom_point() + 
  geom_smooth(data = lms_exp, aes(x = age, y = P5)) + 
  geom_smooth(data = lms_exp, aes(x = age, y = P10)) + 
  geom_smooth(data = lms_exp, aes(x = age, y = P25)) + 
  geom_smooth(data = lms_exp, aes(x = age, y = P50), se = TRUE) + 
  geom_smooth(data = lms_exp, aes(x = age, y = P75)) + 
  geom_smooth(data = lms_exp, aes(x = age, y = P90)) + 
  geom_smooth(data = lms_exp, aes(x = age, y = P95)) 
```

## Growth Plots for Fine Motor Domain
```{r}
# Filter to Cognitive domain
gsv_fm <- gsv %>%
  filter(domain == "Fine Motor")

# Plot Cognitive scores by age
ggplot(gsv_fm, aes(x = bsid_age_calc, y = transformed_score)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  labs(
    title = "Fine Motor Domain Scores vs. Age",
    x = "Age (months)",
    y = "Fine Motor Score"
  )

# Remove NAs and preprocess for modeling
gsv_nona_fm <- gsv_fm %>%
  filter(redcap_event_name != "2_month_visit_arm_1") %>%
  filter(!is.na(bsid_age_calc), !is.na(score))

# Fit GAMLSS model
gamlss_fm <- gamlss(
  formula = transformed_score ~ pb(bsid_age_calc, lambda = 5),
  sigma.formula = ~ pb(bsid_age_calc),
  nu.formula = ~1,
  tau.formula = ~1,
  data = gsv_nona_fm,
  family = BCCG()
)



# 1. Prediction ages
ages <- seq(2, 41, by = 0.5)

# 2. Build newdata with all required covariates used in the model
newdata <- data.frame(bsid_age_calc = ages)

# 3. Get predicted distribution parameters from the model
params <- predictAll(gamlss_fm, newdata = newdata)

# 4. Calculate centiles manually using the BCCG distribution quantile function
centiles <- c(5, 10, 25, 50, 75, 90, 95)
q_vals <- sapply(centiles / 100, function(p) {
  qBCCG(p, mu = params$mu, sigma = params$sigma, nu = params$nu)
})

# 5. Create a dataframe of predicted centiles
lms_fm <- data.frame(age = ages, q_vals)
colnames(lms_fm)[-1] <- paste0("P", centiles)


## Plot it
ggplot(gsv_nona_fm, aes(x = bsid_age_calc, y = transformed_score)) + 
  geom_point() + 
  geom_smooth(data = lms_fm, aes(x = age, y = P5)) + 
  geom_smooth(data = lms_fm, aes(x = age, y = P10)) + 
  geom_smooth(data = lms_fm, aes(x = age, y = P25)) + 
  geom_smooth(data = lms_fm, aes(x = age, y = P50)) + 
  geom_smooth(data = lms_fm, aes(x = age, y = P75)) + 
  geom_smooth(data = lms_fm, aes(x = age, y = P90)) + 
  geom_smooth(data = lms_fm, aes(x = age, y = P95)) 
```

## Growth Plots for Gross Motor Domain
```{r}
# Filter to Cognitive domain
gsv_gm <- gsv %>%
  filter(domain == "Gross Motor")

# Plot Cognitive scores by age
ggplot(gsv_gm, aes(x = bsid_age_calc, y = transformed_score)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  labs(
    title = "Gross Motor Domain Scores vs. Age",
    x = "Age (months)",
    y = "Gross Motor Score"
  )

# Remove NAs and preprocess for modeling
gsv_nona_gm <- gsv_gm %>%
  filter(redcap_event_name != "2_month_visit_arm_1") %>%
  filter(!is.na(bsid_age_calc), !is.na(score))

# Fit GAMLSS model
gamlss_gm <- gamlss(
  formula = transformed_score ~ pb(bsid_age_calc, lambda = 5),
  sigma.formula = ~ pb(bsid_age_calc),
  nu.formula = ~1,
  tau.formula = ~1,
  data = gsv_nona_gm,
  family = BCCG()
)



# 1. Prediction ages
ages <- seq(2, 41, by = 0.5)

# 2. Build newdata with all required covariates used in the model
newdata <- data.frame(bsid_age_calc = ages)

# 3. Get predicted distribution parameters from the model
params <- predictAll(gamlss_gm, newdata = newdata)

# 4. Calculate centiles manually using the BCCG distribution quantile function
centiles <- c(5, 10, 25, 50, 75, 90, 95)
q_vals <- sapply(centiles / 100, function(p) {
  qBCCG(p, mu = params$mu, sigma = params$sigma, nu = params$nu)
})

# 5. Create a dataframe of predicted centiles
lms_gm <- data.frame(age = ages, q_vals)
colnames(lms_gm)[-1] <- paste0("P", centiles)


## Plot it
ggplot(gsv_nona_gm, aes(x = bsid_age_calc, y = transformed_score)) + 
  geom_point() + 
  geom_smooth(data = lms_gm, aes(x = age, y = P5)) + 
  geom_smooth(data = lms_gm, aes(x = age, y = P10)) + 
  geom_smooth(data = lms_gm, aes(x = age, y = P25)) + 
  geom_smooth(data = lms_gm, aes(x = age, y = P50)) + 
  geom_smooth(data = lms_gm, aes(x = age, y = P75)) + 
  geom_smooth(data = lms_gm, aes(x = age, y = P90)) + 
  geom_smooth(data = lms_gm, aes(x = age, y = P95)) 
```

```{r}
view(gsv)
```

```{r}
# Simulate fake reference scores
set.seed(123)

n_ref <- length(composite_long_filtered)
age_range <- range(composite_long$bsid_age_calc, na.rm = TRUE)

ref_data <- data.frame(
  bsid_age_calc = runif(n_ref, min = age_range[1], max = age_range[2]),
  score1 = rnorm(n_ref, mean = 100, sd = 15))

composite_long_filtered_ref <- bind_rows(composite_long_filtered, ref_data)


ggplot(composite_long_filtered, aes(
  x = fct_reorder(composite_long_filtered$sca_condition, composite_long_filtered$bsid_age_calc),
  y = bsid_age_calc,
  fill = sca_condition
)) +
  geom_violin(data = composite_long_filtered, trim = FALSE, alpha = 0.5, color = NA) +
  geom_jitter(data = composite_long_filtered, aes(color = score),
              width = 0.2, height = 0.1, size = 2, alpha = 0.5) + geom_point() + 
geom_smooth(
  data = ref_data,
  aes(x = bsid_age_calc, y = score1),
  method = "lm",
  color = "black",
  se = T,
  linetype = "dashed"
)+
  facet_wrap(~domain, ncol = 5) +
  coord_flip() +
  scale_color_gradient2(low = "blue", mid = "orange", high = "red", midpoint = 100) +
  labs(
    x = NULL,
    y = "Age at Assessment (Months)",
    fill = "SCA Condition",
    color = "Score"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

```

```{r}
ref_data <- expand.grid(
  bsid_age_calc = seq(min(composite_long$bsid_age_calc, na.rm = TRUE),
                      max(composite_long$bsid_age_calc, na.rm = TRUE),
                      length.out = 200),
  sca_condition = unique(composite_long$sca_condition),
  domain = unique(composite_long$domain)  # only if faceting by domain
) %>%
  mutate(
    score = rnorm(n(), mean = 100, sd = 15)
  )

ggplot(composite_long, aes(
  x = fct_reorder(sca_condition, bsid_age_calc),
  y = bsid_age_calc,
  fill = sca_condition
)) +
  geom_violin(trim = FALSE, alpha = 0.5, color = NA) +
  geom_jitter(aes(color = score),
              width = 0.2, height = 0.1, size = 2, alpha = 0.5) +

  # Reference smooth line across score vs. age
  geom_smooth(
    data = ref_data,
    aes(x = bsid_age_calc, y = score, group = sca_condition),
    method = "loess",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    inherit.aes = FALSE
  ) +

  facet_wrap(~domain, ncol = 5) +
  coord_flip() +
  scale_color_gradient2(low = "blue", mid = "orange", high = "red", midpoint = 100) +
  labs(
    x = NULL,
    y = "Age at Assessment (Months)",
    fill = "SCA Condition",
    color = "Score"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

```

```{r}
# Simulate reference data
set.seed(123)
ref_data <- data.frame(
  score1 = rnorm(2478, mean = 100, sd = 15)
)
# view(ref_data)

# composite_combined <- c(composite_long, ref_data)



composite_long_filtered <- composite_long %>% filter(score < 777) %>% mutate(ranorm = rnorm(nrow(composite_long_filtered), mean = 100, sd = 15))

# Plot
ggplot(composite_long_filtered, aes(
  x = fct_reorder(sca_condition, score),
  y = score,
  fill = sca_condition
)) +
  geom_violin(trim = FALSE, alpha = 0.5, color = NA) +
  geom_jitter(aes(color = as.factor(redcap_event_name),
              width = 0.2, height = 0.1, size = 2, alpha = 0.5)) + 

  # Horizontal smooth line (x = age, y = score) -> flips to horizontal under coord_flip
  geom_smooth(
    aes(x = bsid_age_calc, y = ranorm),  # <- This will be flipped by coord_flip()
    method = "loess",
    color = "black",
    se = TRUE,
    linetype = "dashed",
    inherit.aes = FALSE
  ) +

  facet_wrap(~domain, ncol = 5) +
  # scale_color_gradient2(low = "blue", mid = "orange", high = "red", midpoint = 100) +
  labs(
    x = NULL,
    y = "Age at Assessment (Months)",
    fill = "SCA Condition",
    color = "Score"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

```

```{r}
ggplot(composite_long, aes(
  x = score,                # score on x-axis
  y = bsid_age_calc,        # age on y-axis
  fill = sca_condition
)) +
  geom_violin(
    aes(x = sca_condition, y = bsid_age_calc),  # violin by group
    data = composite_long,
    trim = FALSE,
    alpha = 0.5,
    color = NA
  ) +
  geom_jitter(
    aes(color = score),
    width = 0.2, height = 0.1,
    size = 2, alpha = 0.6
  ) +
  # Reference smooth line: horizontal
  geom_smooth(
    data = ref_data,
    aes(x = score, y = bsid_age_calc),
    method = "loess",
    color = "black",
    se = TRUE,
    linetype = "dashed",
    inherit.aes = FALSE
  ) +
  facet_wrap(~domain, ncol = 5) +
  scale_color_gradient2(low = "blue", mid = "orange", high = "red", midpoint = 100) +
  labs(
    x = "Score",
    y = "Age at Assessment (Months)",
    fill = "SCA Condition",
    color = "Score"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

```

